Distribution-oblivious Programming with Dish

### 2019-11-20 pldi/draft.tex
Add a comment about a formal argument in the scaleup section
Minor changes to text

Along with their implicit context---these standard streams and environment variables---commands can express
Command flags (options)

% TODO: Talk about security?
Streams are a key feature of \unix, and go hand-in-hand with the file abstraction:
  files are simply persistent streams that can be manipulated by commands or pipelines.
Reading a file or a set of files generates a stream; 
  similarly, streams can be trivially redirected to the file system---by appending with the file identifier.
Files are named, in the sense that they correspond to a global identifier---for example \ttt{cat /x/y} returns the persistent stream associated with the identifier \ttt{/x/y}.
\unix exposes a few different types of files---\eg directories, symbolic links, named pipes (or FIFOs), and special character devices.
As a result, file-system identifiers part of a pipeline may resolve to in-memory structures or pseudo-devices (\eg \ttt{/dev/urandom}).

@@

Special files are particularly interesting because they are resolved to OS structures.
One example is the proc filesystem (procfs), a special file-system that presents information about processes and other system information.
This file-system enables a form of programmatic introspection, in which commands can inspect their directory or the flags passed to them.
% Here are a few examples:
% (i) \ttt{/proc/PID/cmdline} contains the command that originally started the process,
% (ii) \ttt{/proc/modules} contains a list of the kernel modules currently loaded, and
% (iii) \ttt{/proc/PID/cwd} contains a symlink to the current working directory of the process.
Similarly, \unix exposes file-system handles for various process-related information, include identifiers for ephmeral streams---\ie ones 
OS-internal structures.

There are 3 standard streams (stdin, stdout, stderr) created automatically for most Unix programs.
These correspond to \ttt{/dev/fd/0}.
% These streams are created by the C library (glibc) at the start of program execution,
and new streams can be created to connect to files, sockets, pipes,

This isomorphism is bidirectional---

---and a key enabler of the runtime rebinding achieved during our transformations.
between file-system handles and ephmeral streams

Unix already specifies a few classes of files:

Here are a few special files:
  \ttt{proc} called

Technically, anonymous pipes show up as file identifiers in a hidden filesystem.
Similarly, open files and other ephemeral identifiers

@@

\heading{Stream Manipulation}
A small but powerful set of primitives are concerned with stream manipulation.
First, as described above, several commands can be redirected
Second,

>>>>

The operation of commands can be tuned by two mechanisms, namely command-line options and environment variables.
Options control the execution of a program---\eg changing the \ttt{sort}ing order.
The shell does not have any visibility into these options; 
  their parsing and evaluation is left entirely up to each command.
Environment variables carry more general information about the surrounding system---\eg file search paths, system defaults, parsing switches \etc 
They are organized in a map from names to values, both of which are strings, and can be evaluated at any point in the program---including as commands, flags, a shell feature termed parameter \emph{expansion}.

A small but powerful set of features is concerned with stream manipulation.
Some of these features are built into the shell;
for example, given a file identifier, the stream redirection operator can direct stdout or stderr to the file ---of course, this can be a FIFO that is concurrently used as input by a different command.
More interestingly, some commands allow splitting a single stream into multiple streams while others allow merging streams together (either one after the other, or fusing (zipping) them column-wise into a single stream.

First, as described above, several commands can be redirected
Second,

% TODO: Talk about security primitives?
Streams are a key feature of \unix, and go hand-in-hand with the file abstraction:
  files are simply persistent streams that can be manipulated by commands or pipelines.
Reading a file or a set of files generates a stream; 
  similarly, streams can be trivially redirected to the file system---by appending with the file identifier.
Files are named, in the sense that they correspond to a global identifier---for example \ttt{cat /x/y} returns the persistent stream associated with the identifier \ttt{/x/y}.
\unix exposes a few different types of files---\eg directories, symbolic links, named pipes (or FIFOs), and special character devices.
As a result, file-system identifiers part of a pipeline may resolve to in-memory structures or pseudo-devices (\eg \ttt{/dev/urandom}).

@@

Special files are particularly interesting because they are resolved to OS structures.
One example is the special \ttt{/proc} filesystem (procfs) that presents information about processes and the operating system, offering a feature analogous to programmatic introspection:
  with \ttt{procfs} commands can inspect and alter their directory or the flags passed to them.
% Here are a few examples:
% (i) \ttt{/proc/PID/cmdline} contains the command that originally started the process,
% (ii) \ttt{/proc/modules} contains a list of the kernel modules currently loaded, and
% (iii) \ttt{/proc/PID/cwd} contains a symlink to the current working directory of the process.
Similarly, \unix exposes file-system handles for various ephmeral structures, including standard streams.

The combination of file-system identifiers pointing to ephemeral streams and the stream redirection described above enable surprisingly expressive pipelines---for example:

% These streams are created by the C library (glibc) at the start of program execution,
% and new streams can be created to connect to files, sockets, pipes,

This bidirectional relationship between ephemeral streams and persistent names---the fact that streams can be named in the file-system and files can be converted to streams---is a key enabler of the runtime rebinding achieved during our transformations.

Apart from pipes, the \unix shell features several  other forms of program composition.
Command substitution, enclosed in \ttt{\$()}, replaces the contents of the 
Code block redirects { - }, 
grouping command 
single redirection to redirect the standard output of both commands in the
this is analogous to an anonymous function, as it can be assigned to a variable and be manipulated as a subscript.
A subshell enclosed in \ttt{()} is similar , but executes the block in a new sub-shell---avoiding polluting environment. % but incurring higher overhead.

% \begin{lstlisting}[
%   language=es,
%   mathescape,
%   float=t,
%   belowskip=-5mm,
%   label=core-pl,
%   upquote=true,
%   caption={
%     \textbf{The core language of the POSIX shell.}
%     The language used by ; file identifiers and expansion are not shown.
%   }]
% $p$ ::= $s\in{}String$ | $n\in{}Number$ | $b\in{}Bool$ | $\emptyset$
% $v$ ::= $p$ | ($x$,$\ldots$) => {e} | {$s$:$v$,$\ldots$} | [$v$,$\ldots$]
% $e$ ::= x | $v$ | ($x$ = $e$) $e$ | $e$($e$) | $e$[$e$] = $e$
% \end{lstlisting}

@@
